The Linear Instability of Dilute Ultrarelativistic \epm\ Pair Beams

### 2016-07-01 oblique.tex
fixed subscripts B and b -> b

%\newcommand{\betabar}{\ensuremath{\overline{\boldsymbol \beta}}}
\newcommand{\betabar}{\ensuremath{\overline{\bmath{\beta}}}}
\newcommand{\betabarLAB}{\ensuremath{\betabar_{\rm L}}}
\newcommand{\betabarCM}{\ensuremath{\betabar_{\rm CM}}}
\newcommand{\betaBOOST}{\ensuremath{\beta_{\rm B}}}

@@

The relativistic Maxwellian distribution can be extended to a drifting (or boosted) distribution via an appropriate Lorentz transformation.  The relationship between the energies of the lab (boosted) frame and the COM frame is
\begin{equation}
  E_{\rm CM} = \gamma_B\left(E_{\rm L} - \betaBOOST p_{\rm L,\parallel}\right).
\end{equation}
\cp{\sout{Plugging}Inserting} this into (\ref{eq:relativistic maxwellian}), we find the Maxwell-J{\"u}ttner distribution \citep{1911AnP...340..145J,1975PhRvA..12..686W}
\begin{eqnarray}\label{eq:distribution function beam} %{eq:drifting maxwellian}
%f = \frac{1}{4\pi\gamma_B m^2 T_{\rm L} K_2(m/T_{\rm L})} \exp\left(-\frac {E_{\rm L} - \betaBOOST p_{\rm L,\parallel}} {T_{\rm L}}\right), \\
f = \frac{n_b m_e c^2}{4\pi\gamma_B k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gamma_B(E - \vel_B\cdot \bmath{p})} {k_B T}\right)
\end{eqnarray}
where $K_2$ is the 2nd order modified Bessel function.

@@

%In Appendix~\ref{sec:lorentz}, we explicitly derive the scaling of the
%  parallel and perpendicular velocity dispersions with the Lorentz factor upon
%  boosting the distribution function to the lab frame.
%In the limit in which the perpendicular velocity dispersion in the COM frame is non-relativistic, i.e.,
%$\left|\betaCM\right| \ll 1$, we find to lowest order in $\betaCM$
%\begin{equation}
% \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_{\rm B}^2}.
% \label{eq:betaperp}
%\end{equation}
%This simple scaling of the perpendicular velocity dispersion is an immediate 
%consequence of how transverse velocities transform under Lorentz boosts, and may be understood
%in terms of time dilation.
%
%This result is also anticipated by the standard transformation of temperature under boosts, $T\rightarrow T/\gamma$ (see Equation~\ref{eq:drifting maxwellian}).
%This can be understood by considering $mv^2 \sim kT$ -- we pick up two factors of $1/\gamma$
%from the perpendicular velocity dispersion that is countered by one factor of
%$\gamma$ from the mass scaling. 
%
%For the parallel component of the velocity
%dispersion in the non-relativistic limit, $\left|\betaCM\right| \ll 1$, we find
%to lowest order in $\betaCM$:
%\begin{equation}
% \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_{\rm B}^{4}},
% \label{eq:betapara}
%\end{equation}
%Again this scaling follows immediately from the transformation of velocities under Lorentz boosts, and may be simply understood as a consequence of both time dilation and, for the parallel case, length contraction.
%This scaling of the parallel component of the velocity
%dispersion has important consequences below.
%

@@

% Also plotted in Figures \ref{fig:distribution_non_rel} and \ref{fig:distribution_rel} on the right side is the distribution function that is boosted by $\gamma_B = 10^6$.  Motivated by our discussion in Appendix~\ref{sec:lorentz}, we have scaled the perpendicular component of the velocity by $\gamma_B^{-1}$ and the parallel component by $\gamma_B^{-2}$.

@@

% While these scaling will necessarily break down as $T \rightarrow \gamma_B$, for the problem of interest this is never the case.  Hence, for both the nonrelativistic and mildly relativistic case, the velocity dispersion generally scales like $\gamma_{\rm B}^{-1}$ in the perpendicular direction and like $\gamma_{\rm B}^{-2}$ in the parallel direction provided the boost is much larger than the center-of-mass temperature ($\gamma_B \gg T$).  

@@

%  \caption{Distribution function in velocity space in the center-of-mass frame (left) and the boosted (lab) frame (right) for nonrelativistic temperatures: $T=0.01$ (top) and $0.1$ (bottom).  Note that while we plot $\beta_z$ and $\beta_x$ from -1 to 1, not all values of $\beta_z$ and $\beta_x$ are allowed. \ac{[Can this be shown explicitly, e.g., by making those values NaN, or some other method that results in no plot in that instance?  Also, how are these normalized?]}  On the right plot, we show the distribution in a boosted frame by a Lorentz factor of $10^6$.  As these plots would be compressed near $\beta_z \approx 1$, we have chosen to plot $\beta_z-\beta_B$ on the x-axis.  We have also multiplied the $\beta$ factors by a factor of $\gamma_B^2$ for the z (parallel) component and $\gamma_B$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}. These plots show that a non-relativistic velocity dispersion transforms as $\gamma_B^{-2}$ and $\gamma_B^{-1}$ for the parallel and perpendicular direction, respectively.\label{fig:distribution_non_rel}}

@@

%  \caption{Same as Figure \ref{fig:distribution_non_rel} but for relativistic temperatures: $T=1$ (top) and $10$ (bottom).  Again, we have multiplied the $\beta$ factors on the right plot by a factor of $\gamma_B^2$ for the z (parallel) component and $\gamma_B$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}. Even here the dominant effect of the boost remains the rescaling of the velocity dispersions obtained formally for $T\ll1$.  Nevertheless, higher-order corrections are beginning to spread the distribution function over a velocity range larger than that anticipated by these scalings. 

@@

% Motivated by these scalings, to probe the dependence of the oblique and beam instabilities upon $\gamma_B$ and $T$, we adopt the following simplified distribution function\footnote{This form for $f$ may be explicitly derived in the $T\ll1$ limit by Taylor expanding $\gamma_{\rm B}(E-\beta_{\rm B}p_\parallel)$ around $\bmath{\beta}-\bmath{\beta}_{\rm B}=0$.}: 

@@

%  f = \frac{n_b}{\sqrt{\pi^3\alpha_\parallel\alpha_\perp^2}}\exp\left(-\frac{\left(\beta_{\parallel}-\beta_{\rm B}\right)^2}{\dbetabarpara}-\frac{\beta_{\perp}^2}{\dbetabarperp}\right),

@@

% where $\alpha_\perp\equiv T/\gamma_{\rm B}^2$ and $\alpha_\parallel \equiv T/\gamma_{\rm B}^4 = \alpha_\perp/\gamma_{\rm B}^2$ and here and henceforth we have dropped the subscript L, with all quanties are to be interpreted in the lab frame unless otherwise indicated.

@@

We begin first with the distribution function for the target plasma:
\begin{equation}
 F_t = \left(\frac{1}{2\pi m_e k_B T}\right)^{3/2}\exp\left(-\frac{p^2}{2m_e k_B T}\right),
\end{equation}
where the target plasma is assumed to be nonrelativistic, $\bmath{p}=m_e\vel$ is the nonrelativistic momentum,  and$T$ is the temperature of the background.
For the beam plasma, we adopt the Maxwell-J{\"u}ttner distribution (eq.[\ref{eq:distribution function beam}]).
%\begin{equation}\label{eq:distribution function beam}
%F_b = \frac{n_b}{4\pi\gamma_B m^2 T K_2(m/T_{\rm L})} \exp\left(-\frac {\gammabeam(E - \betabeam \cdot \pmom)} {T}\right),
%\end{equation}
Plugging these into the dispersion relation (\ref{eq:dispersion relation}), we find
%\begin{equation}\label{eq:dispersion kinetic}
% 1 - \frac{\omega_{p,t}^2}{k}\int F_t\frac{k^2 - (\kvec\cdot\betavec)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p - \frac{\omega_{p,t}^2}{k}\int \frac{\kvec\cdot\gradp F_b}{\omega - \kvec\cdot\vel}d^3p  = 0,
%\end{equation}
\begin{equation}\label{eq:dispersion kinetic}
 1 - \frac{\omega_{p,t}^2}{k^2c^2}\int F_t\frac{k^2c^2 - (\kvec\cdot\vel)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p + \frac{m_e\omega_{p,b}^2}{k^2}\int \frac{\kvec\cdot\gradp F_b}{\omega - \kvec\cdot\vel}d^3p  = 0,
\end{equation}
where we have integrated by parts only the second term, associated with target plasma.

We discuss the solution to Equation (\ref{eq:dispersion kinetic}) in Appendix \ref{sec:solution kinetic}.  The associate growth rate for the kinetic beam plasma instability is  
\begin{equation}
\begin{aligned}
\Gamma &\simeq - \Gamma_0
\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{B,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}\\
&\times\left[
\left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
+
\frac{\gamma_B^2 v_{B,x'}^2}{2\cG^2c^2} \left(2 \cG'\mu+ 2\right)
\right]
\exp(-\cG'\mu),\label{eq:kinetic growth rate}
\end{aligned}
\end{equation}
where $\Gamma_0 \equiv \omega_p \gamma_B (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.  This specializes to the beam-plasma growth rate if we take $v_{B,x'} = 0$, which gives:
\begin{equation}
\Gamma_{\rm bp} \simeq - \Gamma_0
\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{B,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
\left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
\exp(-\cG'\mu),\label{eq:bp kinetic growth rate}
\end{equation}

@@

In Figure \ref{fig:ObliqueGMZ}, we plot the growth rate as a function of $\cos\theta = \hat{k}\cdot\vel$, i.e., the angle between the beam and the wave vector, and $k_BT/m_e c^2$.  Here, it is clear that the growth rate reaches it maximum value $\cos\theta \gg 1/\gamma_b$, i.e., at an oblique, but not perpendicular angle.  Moreover the maximum growth rates vary little and are robust for broad range of angles between the wavevector and the beam direction.  Because the maximal oblique growth rate varies little for a wide range of angles and relativistic temperatures for $k_B T/m_ec^2 > 1$, we maximize over $\theta$ in Figure \ref{fig:OGgen} and plot the maximal growth rate as a function of $T$ and $\gamma_B - 1$.  Here for relativistic beams, the maximum growth rate varies little with $T$, varying by less than $10\%$ between hot and cold beams.  Thus, we find: 
\begin{equation}
\Gamma_M
\simeq
\left\{
\begin{aligned}
& 0.38 \Gamma_0 & k_BT/m_ec^2\ll 1\\
& 0.34 \Gamma_0 & k_BT/m_ec^2\gg 1
\end{aligned}
\right.\,,
\end{equation}
for $\gamma_b\gtrsim10$, as seen in Figure \ref{fig:OGgen}.

@@

In Appendix \ref{sec:lorentz}, we estimate how the parallel and perpendicular velocity spreads scale in a Maxwell-Junter distribution.  The relevant results are: 
\begin{equation}
\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {k_B T}{\gamma_{\rm B}^2 m_e c^2 },
\end{equation}
and 
\begin{equation}
\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T}{\gamma_{\rm B}^4 m_e c^2 },
\end{equation}
where $T$ is measured in the COM frame of the beam.

@@

Now let us consider moments of the distribution function.  For clarity, it is helpful to consider moments of the distribution function first in the COM frame.  The velocity moment is (unsurprisingly):
\begin{equation}
 \betabarCM = \int \frac {\pmom_{\rm CM}}{\gamma_{\rm CM} mc} \fCM \dVCM = 0.
\end{equation}
We consider the lab frame to be  boosted along the x-axis by $\betaBOOST$. More precisely, the initial inertial frame is the COM frame and the lab frame is moving with velocity $\betaBOOST = -|\betaBOOST |$ with respect to the COM frame. This gives:
\begin{eqnarray}\label{eq:betabarLAB}
 \betabarLAB &=& \int \frac {\pmom_{\rm L}}{\gamma_{\rm L} mc} \fLAB \dVLAB  \nonumber\\
&=& \int \frac {\pmom_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM})}{\gamma_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM}) mc} \fCM \dVCM.
\end{eqnarray}
Breaking the components of \betabarLAB\ into components parallel and
perpendicular to the boost, we find: 
\begin{eqnarray}
\betabarLABpara &=& \int \frac{\betaCMpara -\betaBOOST}{1-\betaBOOST\betaCMpara} \fCM \dVCM,\\
\betabarLABperp &=& \int \frac{\betaCMperp}{\gamma_{\rm B}\left(1-\betaBOOST\betaCMpara\right)} \fCM \dVCM,
\end{eqnarray}
where $\gamma_{\rm B} = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
For $\left|\betaCM\right|,\betaBOOST \ll 1$, we recover the Galilean invariant result, $\betabarLAB \approx \betabarCM - \betaBOOST\hat{\boldsymbol x}$.  
However, this Galilean result no longer holds for relativistic motion.

Now we consider the dispersion around $\betabar$.  In components, the COM frame is:
\begin{equation}
 \dbetabarCMi = \int \beta_i^2 \fCM \dVCM.
\end{equation}
In the lab frame, it is again useful to break it into components -- the parallel component becomes 
\begin{eqnarray}
  \dbetabarLABpara &=& \int (\beta_{\rm L, \parallel}-\betabarLABpara)^2 \fLAB \dVLAB ,\nonumber\\ 
 &=& \int \beta_{\rm L,\parallel}^2 \fLAB \dVLAB - \betabarLABpara^2,\nonumber\\
 &=&\int \left(\frac{\betaCMpara -\betaBOOST}{1-\betaMEAN^2}\right)^2 \fCM \dVCM - \betabarLABpara^2,\label{eq:dbetabarLABpara}
\end{eqnarray}
while the perpendicular component becomes
\begin{eqnarray}
 \dbetabarLABperp &=& \int \beta_{\rm L,\perp}^2 \fLAB \dVLAB, \nonumber\\
&=& \gamma_{\rm B}^{-2} \int \frac{\betaCMperp^2}{\left(1-\betaMEAN^2\right)^2} \fCM \dVCM,
\label{eq:dbetabarLABperp}
\end{eqnarray}
where we have defined $\betaMEAN^2 = \betaCMpara\betaBOOST$ as the geometric mean for matters of convenience.

It is easier to look at the perpendicular component first.  For $\left|\betaCM\right| \ll 1$, Equation (\ref{eq:dbetabarLABperp}) becomes to lowest order in $\betaCM$ 
\begin{equation}
 \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_{\rm B}^2} \approx \frac {k_B T}{\gamma_{\rm B}^2 m_e c^2 }.
\end{equation}
This simple scaling of the perpendicular velocity dispersion can be understood as a scaling with time between two frames boosted relative to each other, where the coordinates perpendicular to the boost axis remain invariant.  This result is also in line with the transformation of temperature as $T\rightarrow T/\gamma$ under a boost, i.e., $mv^2 \sim kT$ -- two factors of $1/\gamma$ from the perpendicular velocity dispersion is countered by one factor of $\gamma$ from the mass.  Let us now consider the parallel component (eq.(\ref{eq:dbetabarLABpara})) again to lowest order in $\betaCM$: 
\begin{equation}
 \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_{\rm B}^{4}}  \approx \frac {k_B T}{\gamma_{\rm B}^4 m_e c^2 },
\end{equation}
Here, the scaling of the parallel velocity dispersion can be understood as a double scaling of both time and coordinate (along the boost axis) between same two frames boosted relative to each other, giving an extra scaling of $\gamma^{-2}$.
This scaling of the parallel component of the velocity dispersion has important consequences that we explore in the main part of the paper.

@@

Putting this all together, the residue is
\begin{equation}
R
=
\frac{
  \gph^3 (\vph-v_{B,z'})\mu^2}{4\pi K_2(\mu)
}
\frac{\kI}{m_e^4c^7}
,
\label{eq:Rogen}
\end{equation}
where
\begin{equation}\label{eq:I-integral}
\kI \equiv \int d^2\!p_\perp E_\perp \exp\left(-\frac{\cG(E_\perp-w p_x)}{k_B T}\right)
\end{equation}
and $\cG\equiv\gamma_B\gph(1-v_{B,z'}\vph/c^2)$ and
$w\equiv \gamma_Bv_{B,x'}/\cG \le 1$.  This latter inequality is guaranteed as 
\begin{equation}
\cG E_\perp - \gamma_B v_{B,x'} p_{x'} = \gamma_B(E - v_{B,z'} p_{z'} - v_{B,x'}p_{x'})|_{v_{B,z'}=v_{\rm ph}} > 0
\end{equation}
is the energy in beam frame and is therefore positive definite.  Noting that the $\exp(-wp_{x'})$ term appears as a boosted distribution, we boost by $w$
along the $x'$-axis, removing the anisotropic term from the
exponential.

>>>>

%\newcommand{\betabar}{\ensuremath{\overline{\boldsymbol \beta}}}
\newcommand{\betabar}{\ensuremath{\overline{\bmath{\beta}}}}
\newcommand{\betabarLAB}{\ensuremath{\betabar_{\rm L}}}
\newcommand{\betabarCM}{\ensuremath{\betabar_{\rm CM}}}
\newcommand{\betaBOOST}{\ensuremath{\beta_b}}

@@

The relativistic Maxwellian distribution can be extended to a drifting (or boosted) distribution via an appropriate Lorentz transformation.  The relationship between the energies of the lab (boosted) frame and the COM frame is
\begin{equation}
  E_{\rm CM} = \gamma_b\left(E_{\rm L} - \betaBOOST p_{\rm L,\parallel}\right).
\end{equation}
\cp{\sout{Plugging}Inserting} this into (\ref{eq:relativistic maxwellian}), we find the Maxwell-J{\"u}ttner distribution \citep{1911AnP...340..145J,1975PhRvA..12..686W}
\begin{eqnarray}\label{eq:distribution function beam} %{eq:drifting maxwellian}
%f = \frac{1}{4\pi\gamma_b m^2 T_{\rm L} K_2(m/T_{\rm L})} \exp\left(-\frac {E_{\rm L} - \betaBOOST p_{\rm L,\parallel}} {T_{\rm L}}\right), \\
f = \frac{n_b m_e c^2}{4\pi\gamma_b k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gamma_b(E - \vel_b\cdot \bmath{p})} {k_B T}\right)
\end{eqnarray}
where $K_2$ is the 2nd order modified Bessel function.

@@

%In Appendix~\ref{sec:lorentz}, we explicitly derive the scaling of the
%  parallel and perpendicular velocity dispersions with the Lorentz factor upon
%  boosting the distribution function to the lab frame.
%In the limit in which the perpendicular velocity dispersion in the COM frame is non-relativistic, i.e.,
%$\left|\betaCM\right| \ll 1$, we find to lowest order in $\betaCM$
%\begin{equation}
% \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_b^2}.
% \label{eq:betaperp}
%\end{equation}
%This simple scaling of the perpendicular velocity dispersion is an immediate 
%consequence of how transverse velocities transform under Lorentz boosts, and may be understood
%in terms of time dilation.
%
%This result is also anticipated by the standard transformation of temperature under boosts, $T\rightarrow T/\gamma$ (see Equation~\ref{eq:drifting maxwellian}).
%This can be understood by considering $mv^2 \sim kT$ -- we pick up two factors of $1/\gamma$
%from the perpendicular velocity dispersion that is countered by one factor of
%$\gamma$ from the mass scaling. 
%
%For the parallel component of the velocity
%dispersion in the non-relativistic limit, $\left|\betaCM\right| \ll 1$, we find
%to lowest order in $\betaCM$:
%\begin{equation}
% \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_b^{4}},
% \label{eq:betapara}
%\end{equation}
%Again this scaling follows immediately from the transformation of velocities under Lorentz boosts, and may be simply understood as a consequence of both time dilation and, for the parallel case, length contraction.
%This scaling of the parallel component of the velocity
%dispersion has important consequences below.
%

@@

% Also plotted in Figures \ref{fig:distribution_non_rel} and \ref{fig:distribution_rel} on the right side is the distribution function that is boosted by $\gamma_b = 10^6$.  Motivated by our discussion in Appendix~\ref{sec:lorentz}, we have scaled the perpendicular component of the velocity by $\gamma_b^{-1}$ and the parallel component by $\gamma_b^{-2}$.

@@

% While these scaling will necessarily break down as $T \rightarrow \gamma_b$, for the problem of interest this is never the case.  Hence, for both the nonrelativistic and mildly relativistic case, the velocity dispersion generally scales like $\gamma_b^{-1}$ in the perpendicular direction and like $\gamma_b^{-2}$ in the parallel direction provided the boost is much larger than the center-of-mass temperature ($\gamma_b \gg T$).  

@@

%  \caption{Distribution function in velocity space in the center-of-mass frame (left) and the boosted (lab) frame (right) for nonrelativistic temperatures: $T=0.01$ (top) and $0.1$ (bottom).  Note that while we plot $\beta_z$ and $\beta_x$ from -1 to 1, not all values of $\beta_z$ and $\beta_x$ are allowed. \ac{[Can this be shown explicitly, e.g., by making those values NaN, or some other method that results in no plot in that instance?  Also, how are these normalized?]}  On the right plot, we show the distribution in a boosted frame by a Lorentz factor of $10^6$.  As these plots would be compressed near $\beta_z \approx 1$, we have chosen to plot $\beta_z-\beta_b$ on the x-axis.  We have also multiplied the $\beta$ factors by a factor of $\gamma_b^2$ for the z (parallel) component and $\gamma_b$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}. These plots show that a non-relativistic velocity dispersion transforms as $\gamma_b^{-2}$ and $\gamma_b^{-1}$ for the parallel and perpendicular direction, respectively.\label{fig:distribution_non_rel}}

@@

%  \caption{Same as Figure \ref{fig:distribution_non_rel} but for relativistic temperatures: $T=1$ (top) and $10$ (bottom).  Again, we have multiplied the $\beta$ factors on the right plot by a factor of $\gamma_b^2$ for the z (parallel) component and $\gamma_b$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}. Even here the dominant effect of the boost remains the rescaling of the velocity dispersions obtained formally for $T\ll1$.  Nevertheless, higher-order corrections are beginning to spread the distribution function over a velocity range larger than that anticipated by these scalings. 

@@

% Motivated by these scalings, to probe the dependence of the oblique and beam instabilities upon $\gamma_b$ and $T$, we adopt the following simplified distribution function\footnote{This form for $f$ may be explicitly derived in the $T\ll1$ limit by Taylor expanding $\gamma_b(E-\beta_b p_\parallel)$ around $\bmath{\beta}-\bmath{\beta}_{\rm B}=0$.}: 

@@

%  f = \frac{n_b}{\sqrt{\pi^3\alpha_\parallel\alpha_\perp^2}}\exp\left(-\frac{\left(\beta_{\parallel}-\beta_b\right)^2}{\dbetabarpara}-\frac{\beta_{\perp}^2}{\dbetabarperp}\right),

@@

% where $\alpha_\perp\equiv T/\gamma_b^2$ and $\alpha_\parallel \equiv T/\gamma_b^4 = \alpha_\perp/\gamma_b^2$ and here and henceforth we have dropped the subscript L, with all quanties are to be interpreted in the lab frame unless otherwise indicated.

@@

We begin first with the distribution function for the target plasma:
\begin{equation}
 F_t = \left(\frac{1}{2\pi m_e k_B T}\right)^{3/2}\exp\left(-\frac{p^2}{2m_e k_B T}\right),
\end{equation}
where the target plasma is assumed to be nonrelativistic, $\bmath{p}=m_e\vel$ is the nonrelativistic momentum,  and$T$ is the temperature of the background.
For the beam plasma, we adopt the Maxwell-J{\"u}ttner distribution (eq.[\ref{eq:distribution function beam}]).
%\begin{equation}\label{eq:distribution function beam}
%F_b = \frac{n_b}{4\pi\gamma_b m^2 T K_2(m/T_{\rm L})} \exp\left(-\frac {\gammabeam(E - \betabeam \cdot \pmom)} {T}\right),
%\end{equation}
Plugging these into the dispersion relation (\ref{eq:dispersion relation}), we find
%\begin{equation}\label{eq:dispersion kinetic}
% 1 - \frac{\omega_{p,t}^2}{k}\int F_t\frac{k^2 - (\kvec\cdot\betavec)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p - \frac{\omega_{p,t}^2}{k}\int \frac{\kvec\cdot\gradp F_b}{\omega - \kvec\cdot\vel}d^3p  = 0,
%\end{equation}
\begin{equation}\label{eq:dispersion kinetic}
 1 - \frac{\omega_{p,t}^2}{k^2c^2}\int F_t\frac{k^2c^2 - (\kvec\cdot\vel)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p + \frac{m_e\omega_{p,b}^2}{k^2}\int \frac{\kvec\cdot\gradp F_b}{\omega - \kvec\cdot\vel}d^3p  = 0,
\end{equation}
where we have integrated by parts only the second term, associated with target plasma.

We discuss the solution to Equation (\ref{eq:dispersion kinetic}) in Appendix \ref{sec:solution kinetic}.  The associate growth rate for the kinetic beam plasma instability is  
\begin{equation}
\begin{aligned}
\Gamma &\simeq - \Gamma_0
\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}\\
&\times\left[
\left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
+
\frac{\gamma_b^2 v_{b,x'}^2}{2\cG^2c^2} \left(2 \cG'\mu+ 2\right)
\right]
\exp(-\cG'\mu),\label{eq:kinetic growth rate}
\end{aligned}
\end{equation}
where $\Gamma_0 \equiv \omega_p \gamma_b (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.  This specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
\begin{equation}
\Gamma_{\rm bp} \simeq - \Gamma_0
\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
\left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
\exp(-\cG'\mu),\label{eq:bp kinetic growth rate}
\end{equation}

@@

In Figure \ref{fig:ObliqueGMZ}, we plot the growth rate as a function of $\cos\theta = \hat{k}\cdot\vel$, i.e., the angle between the beam and the wave vector, and $k_BT/m_e c^2$.  Here, it is clear that the growth rate reaches it maximum value $\cos\theta \gg 1/\gamma_b$, i.e., at an oblique, but not perpendicular angle.  Moreover the maximum growth rates vary little and are robust for broad range of angles between the wavevector and the beam direction.  Because the maximal oblique growth rate varies little for a wide range of angles and relativistic temperatures for $k_B T/m_ec^2 > 1$, we maximize over $\theta$ in Figure \ref{fig:OGgen} and plot the maximal growth rate as a function of $T$ and $\gamma_b - 1$.  Here for relativistic beams, the maximum growth rate varies little with $T$, varying by less than $10\%$ between hot and cold beams.  Thus, we find: 
\begin{equation}
\Gamma_M
\simeq
\left\{
\begin{aligned}
& 0.38 \Gamma_0 & k_BT/m_ec^2\ll 1\\
& 0.34 \Gamma_0 & k_BT/m_ec^2\gg 1
\end{aligned}
\right.\,,
\end{equation}
for $\gamma_b\gtrsim10$, as seen in Figure \ref{fig:OGgen}.

@@

In Appendix \ref{sec:lorentz}, we estimate how the parallel and perpendicular velocity spreads scale in a Maxwell-Junter distribution.  The relevant results are: 
\begin{equation}
\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {k_B T}{\gamma_b^2 m_e c^2 },
\end{equation}
and 
\begin{equation}
\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T}{\gamma_b^4 m_e c^2 },
\end{equation}
where $T$ is measured in the COM frame of the beam.

@@

Now let us consider moments of the distribution function.  For clarity, it is helpful to consider moments of the distribution function first in the COM frame.  The velocity moment is (unsurprisingly):
\begin{equation}
 \betabarCM = \int \frac {\pmom_{\rm CM}}{\gamma_{\rm CM} mc} \fCM \dVCM = 0.
\end{equation}
We consider the lab frame to be  boosted along the x-axis by $\betaBOOST$. More precisely, the initial inertial frame is the COM frame and the lab frame is moving with velocity $\betaBOOST = -|\betaBOOST |$ with respect to the COM frame. This gives:
\begin{eqnarray}\label{eq:betabarLAB}
 \betabarLAB &=& \int \frac {\pmom_{\rm L}}{\gamma_{\rm L} mc} \fLAB \dVLAB  \nonumber\\
&=& \int \frac {\pmom_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM})}{\gamma_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM}) mc} \fCM \dVCM.
\end{eqnarray}
Breaking the components of \betabarLAB\ into components parallel and
perpendicular to the boost, we find: 
\begin{eqnarray}
\betabarLABpara &=& \int \frac{\betaCMpara -\betaBOOST}{1-\betaBOOST\betaCMpara} \fCM \dVCM,\\
\betabarLABperp &=& \int \frac{\betaCMperp}{\gamma_b\left(1-\betaBOOST\betaCMpara\right)} \fCM \dVCM,
\end{eqnarray}
where $\gamma_b = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
For $\left|\betaCM\right|,\betaBOOST \ll 1$, we recover the Galilean invariant result, $\betabarLAB \approx \betabarCM - \betaBOOST\hat{\boldsymbol x}$.  
However, this Galilean result no longer holds for relativistic motion.

Now we consider the dispersion around $\betabar$.  In components, the COM frame is:
\begin{equation}
 \dbetabarCMi = \int \beta_i^2 \fCM \dVCM.
\end{equation}
In the lab frame, it is again useful to break it into components -- the parallel component becomes 
\begin{eqnarray}
  \dbetabarLABpara &=& \int (\beta_{\rm L, \parallel}-\betabarLABpara)^2 \fLAB \dVLAB ,\nonumber\\ 
 &=& \int \beta_{\rm L,\parallel}^2 \fLAB \dVLAB - \betabarLABpara^2,\nonumber\\
 &=&\int \left(\frac{\betaCMpara -\betaBOOST}{1-\betaMEAN^2}\right)^2 \fCM \dVCM - \betabarLABpara^2,\label{eq:dbetabarLABpara}
\end{eqnarray}
while the perpendicular component becomes
\begin{eqnarray}
 \dbetabarLABperp &=& \int \beta_{\rm L,\perp}^2 \fLAB \dVLAB, \nonumber\\
&=& \gamma_b^{-2} \int \frac{\betaCMperp^2}{\left(1-\betaMEAN^2\right)^2} \fCM \dVCM,
\label{eq:dbetabarLABperp}
\end{eqnarray}
where we have defined $\betaMEAN^2 = \betaCMpara\betaBOOST$ as the geometric mean for matters of convenience.

It is easier to look at the perpendicular component first.  For $\left|\betaCM\right| \ll 1$, Equation (\ref{eq:dbetabarLABperp}) becomes to lowest order in $\betaCM$ 
\begin{equation}
 \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_b^2} \approx \frac {k_B T}{\gamma_b^2 m_e c^2 }.
\end{equation}
This simple scaling of the perpendicular velocity dispersion can be understood as a scaling with time between two frames boosted relative to each other, where the coordinates perpendicular to the boost axis remain invariant.  This result is also in line with the transformation of temperature as $T\rightarrow T/\gamma$ under a boost, i.e., $mv^2 \sim kT$ -- two factors of $1/\gamma$ from the perpendicular velocity dispersion is countered by one factor of $\gamma$ from the mass.  Let us now consider the parallel component (eq.(\ref{eq:dbetabarLABpara})) again to lowest order in $\betaCM$: 
\begin{equation}
 \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_b^{4}}  \approx \frac {k_B T}{\gamma_b^4 m_e c^2 },
\end{equation}
Here, the scaling of the parallel velocity dispersion can be understood as a double scaling of both time and coordinate (along the boost axis) between same two frames boosted relative to each other, giving an extra scaling of $\gamma^{-2}$.
This scaling of the parallel component of the velocity dispersion has important consequences that we explore in the main part of the paper.

@@

Putting this all together, the residue is
\begin{equation}
R
=
\frac{
  \gph^3 (\vph-v_{b,z'})\mu^2}{4\pi K_2(\mu)
}
\frac{\kI}{m_e^4c^7}
,
\label{eq:Rogen}
\end{equation}
where
\begin{equation}\label{eq:I-integral}
\kI \equiv \int d^2\!p_\perp E_\perp \exp\left(-\frac{\cG(E_\perp-w p_x)}{k_B T}\right)
\end{equation}
and $\cG\equiv\gamma_b\gph(1-v_{b,z'}\vph/c^2)$ and
$w\equiv \gamma_bv_{b,x'}/\cG \le 1$.  This latter inequality is guaranteed as 
\begin{equation}
\cG E_\perp - \gamma_b v_{b,x'} p_{x'} = \gamma_b(E - v_{b,z'} p_{z'} - v_{b,x'}p_{x'})|_{v_{b,z'}=v_{\rm ph}} > 0
\end{equation}
is the energy in beam frame and is therefore positive definite.  Noting that the $\exp(-wp_{x'})$ term appears as a boosted distribution, we boost by $w$
along the $x'$-axis, removing the anisotropic term from the
exponential.
